DynaProg for Scala

### 2013-01-02 docs/report/design.tex
Nits.

% challenge in programming CUDA devices is to make efficient 
{\color{red} XXX}

>>>>

\begin{figure}[H]\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(16,2.5)
	\put(1,0){\moveto(2,0)\lineto(2,2)\lineto(0,2)\closepath\strokepath\put(1.1,1.1){\linethickness{1.5pt}\vector(1,1){.75}}
		\multiput(0,0)(0,.1){20}{\circle*{.03}}\multiput(0,0)(.1,0){20}{\circle*{.03}}}
	\put(0,-.8){\makebox[4\unitlength]{Triangle}}

	\put(7,0){\moveto(0,0)\lineto(2,0)\lineto(2,2)\lineto(0,2)\closepath\strokepath
		\put(.5,1.5){\linethickness{1.5pt}\vector(1,-1){1}}}
	\put(6,-.8){\makebox[4\unitlength]{Rectangle}}

	\put(12,0){\moveto(2,0)\lineto(4,0)\lineto(2,2)\lineto(0,2)\closepath\strokepath
		\put(1.6,0.6){\linethickness{1.5pt}\vector(1,1){.8}}
		\multiput(0,0)(2,0){3}{\multiput(0,0)(0,.1){20}{\circle*{.03}}}
		\multiput(0,0)(.1,0){20}{\circle*{.03}}\multiput(4,2)(-.1,0){20}{\circle*{.03}}}
	\put(12,-.8){\makebox[4\unitlength]{Parallelogram}}
\end{picture}\end{center}\caption{Matrix shapes, the arrow indicates the computation progress direction}\end{figure}

Since progress is made along the diagonal of the matrices, a naive row/column addressing of the matrix elements would result in no coalesced access. Hence instead of addressing by row the matrix elements, we address them by diagonal. Let $M_W$ the width of the matrix and $M_H$ its height. Coalesced addressing is most easy in the parallelogram matrix, as we just need to map parallelogram indices to addresses into a square matrix (with diagonal elements being contiguous in memory) as follows:
\[(i,j) \to i + (j-i)*M_H\]

Noticing that the triangle is just half the parallelogram, we could use the same addressing, but we would use twice the memory necessary. Hence we need a more involved formula: the size of the triangle embedded in a square of side $k$ (incl. diagonal) is $\frac{k\cdot(k+1)}{2}$ knowing that fact, the index of the $i^{\rm th}$ element on the $j^{\rm th}$ diagonal of the triangle (starting with the main diagonal when $j=0$), we obtain the following formula (with $M_H=M_W=n$):
\[(i,j) \to M-T+i \text{ with }\left\{\begin{array}{rcll}
	M&=&\dfrac{(n\cdot (n+1)}{2} \vspace{.15cm} & \textit{total }\bigtriangleup\\
	d&=&n+1+i-j \vspace{.15cm} & \textit{diagonal of }(i,j)\\
	T &=&\dfrac{(d\cdot (d+1)}{2} & \bigtriangleup \textit{ of current diagonal}
\end{array}\right.\]

Finally, for the rectangular matrix, since the parallelogram indexing looks efficient, we want to reuse the same idea. However, embedding the rectangle within a parallelogram would have a very large overhead ($(M_H)^2$, by adding a triangle on each side of the rectangle). A solution consists into breaking the rectangle into multiple horizontal stripes of fixed height $B_H$, hereby dramatically reducing the size of the triangles needed. Finally, the stripes can be stitched together to form a single parallelogram (beyond a certain number of lines, coalescing access does not improve as memory is anyway not stored contiguously).

\begin{figure}[H]\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(4,3.5)
\put(0,-.5) {
	\put(0,0){\color{lightgray}\moveto(4,0)\lineto(4.5,.5)\lineto(4,.5)\closepath\fillpath\moveto(0,4)\lineto(-.5,3.5)\lineto(0,3.5)\closepath\fillpath}
	\put(1,3){\linethickness{1.5pt}\vector(1,-1){2}}
	\multiput(0,.5)(0,.5){7}{\line(1,0){4}}
	\moveto(0,0)\lineto(4,0)\lineto(4.5,.5)\lineto(4,.5)\lineto(4,4)\lineto(0,4)\lineto(-.5,3.5)\lineto(0,3.5)\closepath\strokepath
}
\end{picture}\end{center}\caption{Parallelogram memory representation of a rectangular matrix}\end{figure}

It follows that the total memory required to store the matrix is:
\[M_W \cdot \left\lceil\dfrac{M_H}{B_H}\right\rceil \cdot B_H  + (B_H)^2\]

and the mapping of indices is given by
\[ (i,j) \to (B_H \cdot \underset{\rm diagonal}{(j + m)} + m + \underset{\text{stripe}}{\left\lfloor \dfrac{i}{B_H} \right\rfloor} \cdot \underset{\text{stripe size}}{M_W \cdot B_H}) \quad \text{with } m=i \mod B_H \]
